---
title: "GLHLTH562 Final Project"
subtitle: "<br/>Income Tax Revenue Projections"
author: "Sam Sloate"
institute: "Duke University"
date: "4/13/2022"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css,custom_deck.css]
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 4:3
      countIncrementalSlides: true
    seal: false
---

class: titleSlide 

```{r xsetup, include=FALSE}
library(tidyverse)
library(xaringan)
library(xaringanthemer)
library(xaringanExtra)
library(hrbrthemes)
library(knitr)
library(emo)

import_roboto_condensed()

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.width=9, fig.height=3.5, fig.retina=3, out.width = "100%", cache = FALSE, echo = TRUE, message = FALSE, warning = FALSE, hiline = TRUE)

# Bring in files
purl(here::here("final.Rmd"))
source(here::here("final.R"))

purl(here::here("sensitivity_analysis.rmd"))
source(here::here("sensitivity_analysis.R"))

# Assign color scheme
blue1<- "#6a92b9"
blue2 <- "#0B3B60"

gray1 <- "#82858c"
gray2 <- "#d4d7db"

```


```{r xaringan-themer, include=FALSE, warning=FALSE}
style_duo_accent(
  primary_color = "#0B3B60",
  secondary_color = "#6a92b9",
  header_font_google = google_font("Roboto Condensed", font_family="Bold"),
  text_font_google   = google_font("Roboto Condensed", font_family="Light"),
  code_font_google   = google_font("Anonymous Pro"),
  base_font_size = "30px",
  text_font_size = "1rem",
  header_h1_font_size = "3.5rem",
  header_h2_font_size = "2rem",
  header_h3_font_size = "1rem",
  padding = "8px 32px 8px 32px",
)
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://malek.icourban.com/crypto-https-files.nc.gov/osbm/OSBM-logo-header.png?VersionId=Kxo6NFAKpIfvLfY5GNU5oWjHE8NAyj4f"
)
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$subtitle`

<br>

`r emo::ji("graph")`
---
class: center, middle

# Objective

Help the Office of State Budget and Management <br> understand **changes in demographics** <br> and **future income tax liability**.
---
class: left

# Research Questions
.pull-left[
### 1 
How are North Carolina demographic trends projected to change over time?
]
.pull-right[
###2 
How will these changes affect North Carolina’s income tax revenue in 2040?
]

---
class: newTopic, hide_logo

# Demographic Changes

---

class: left

```{r echo=FALSE, fig.height=6, fig.width=8, fig.align="center"}
popplot
```

---

class: left
```{r echo=FALSE, fig.height=6, fig.width=8, fig.align="center"}
ageplot
```

---

class: left
```{r echo=FALSE, fig.height=6, fig.width=8, fig.align="center"}
raceplot
```

---
class: newTopic, hide_logo

# Model Methodology
---

class: left

## Personal Income Tax Microsimulation Model
* Uses Current Population Survey (CPS) microdata and OSBM demographic projections

* Adjusts the weights of each observation to simulate changes in the projected number of people with a given set of characteristics

* Adjusts incomes over time

* Calculates North Carolina tax liability 

---

class: newTopic, hide_logo

# Model Results

---

class: left

```{r echo=FALSE, fig.height=6, fig.width=8,  fig.align="center"}
taxliab_plot
```

---

class: left

## Tax Liability

<br>
<br>

What is the main driver of increasing tax liability?
<br>
*Population? Income Growth?*
---

class: left

```{r echo=FALSE, fig.height=6, fig.width=8,  fig.align="center"}
noincgrowth_plot
```

---

class: newTopic, hide_logo

# Sensitivity Analyses

---

class: left, hide-count

## Sensitivity Analyses

.panelset[
.panel[.panel-name[Scenario 1]

**Slower and Faster Income Growth**

* +50% growth rates and 3% inflation

* -50% growth rates and 1% inflation
]

.panel[.panel-name[Scenario 2]

**Stable Tax Rate**

* Current tax rate: 5.25%

* Current legislation will lower it to 3.99% by 2027

* Scenario 2 keeps the tax rate at **current levels**
]

.panel[.panel-name[Scenario 3]

**Stable Standard Deduction**

* Current standard deduction is between $10,750 (single) and $21,500 (married)

* The model assumes this will grow a set amount per year to roughly double by 2040

* Scenario 3 keeps the standard deduction at **current levels**
]

.panel[.panel-name[Scenario 4]

**No Demographic Shifts**

* Current demographics show percentages of White, Black, and young residents declining over time

* Scenario 4 increases the population to 2040 levels but increases all age and race groups proportionally

* **Do demographic shifts matter?**
]

]

---

class: left

### Scenario 1: Income Growth Changes
```{r echo=FALSE, fig.height=6, out.width="100%",  fig.align="center"}
upper_lower_plot
```

---

class: left

### Scenario 2: Stable Tax Rate
```{r echo=FALSE,  fig.height=6, out.width="100%",   fig.align="center"}
taxrate_plot
```

---

class: left

### Scenario 3: Stable Standard Deduction
```{r echo=FALSE, fig.height=6, out.width="100%",   fig.align="center"}
stdded_plot
```

---

class: left

### Scenario 4: No Demographic Shifts
```{r echo=FALSE, fig.height=6, out.width="100%",   fig.align="center"}
nodemchange_plot
```

---

class: newTopic, hide_logo

# Conclusion

---

class: left

# Takeaways
* 2040 income tax liability is predicted to be **$19.8B**

---

class: left

# Takeaways
* 2040 income tax liability is predicted to be **$19.8B**

* **Income growth** is the main driver of tax liability

---
class: left

# Takeaways
* 2040 income tax liability is predicted to be **$19.8B**

* **Income growth** is the main driver of tax liability

* Demographic shifts are **not** as important for predicting tax liability

---

class: newTopic 

# Thanks!

`r emo::ji("graph")`



